\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{graphicx,subfigure}
\usepackage{enumerate}
\usepackage{hyperref}
\usepackage{amsmath, mathtools, amssymb, amsthm}
\usepackage[margin=0.7in]{geometry}
\usepackage{soul}

\title{Entropy Stability}
\author{(Your name here)}
\date{February 2nd, 2023}

\begin{document}
\maketitle


Dig the literature to provide answers to the following questions. \underline{You cannot refer to any paper published after 1985}. You can read more recent papers - but they should be pointing to earlier works.
\begin{enumerate}
    \item Who came up with the entropy condition? Oldest readable reference you can find. Where does that condition come from? What are the motivations?
    \item Find two first-order numerical fluxes that do satisfy this condition, under certain assumptions. Success stories (did they bring tangible improvements)? There are more than two.
    \item Philip Roe is a legend in CFD, in part for his 'Roe flux'. 
    \begin{enumerate}
        \item Find the paper where he first introduced it.
        \item How did Phil Roe derive the 'Roe average'? What was he after that led to it?
    \end{enumerate}    
    \item One approach to entropy-stability was introduced by Tadmor in 1987 ( read \cite{Tadmor1987}). What distinguishes it from others?
    \item It took quite some time for Tadmor's ideas to gain popularity. Moon mentioned work by Fisher and Carpenter in 2013, where robustness potential for higher-order schemes was demonstrated for the first time. But Phil Roe made key contributions in 2006 (see \cite{Roe2006}). Discuss them
    \begin{itemize}
        \item[-] Why is there a logarithmic average in his flux?
    \end{itemize}
\end{enumerate}
Consider the 1D compressible Euler equations for a perfect gas,
\begin{enumerate}
    \item Derive the conservation equation for entropy $(\rho s)$. 
    \item How would you formalize the derivation you just carried out? Say you have a system of conservation laws with $\mathbf{u}$ and $\mathbf{f}$, and you are asked to derive the conservation equation (assuming there is one) for a scalar valued function $U = U(\mathbf{u})$.
    \item Does the number of spatial dimensions matter?
    \item Is $(\rho s)$ a convex function of mass, momentum and total energy? Define convexity first.
    \item What are numerical and theoretical benefits of convexity?
    \item Prove that the original Roe flux is not entropy-stable, using Tadmor's framework \cite{Tadmor1987} and a code implementation of the Roe flux.  
    \begin{itemize}
        \item People were already aware that the Roe flux was not entropy-stable. A correction called 'entropy fix' was introduced. Find the paper that first introduced it. 
        \item Is the Roe flux with the entropy-fix entropy-stable ?  
    \end{itemize}
\end{enumerate}

\begin{thebibliography}{9}
\input{bibliography.tex}
\end{thebibliography}
\end{document}
